---
title: "Multiple Linear Regression"
subtitle: "Lecture 6"
date: "2026-09-17"
format: 
  revealjs: 
    output-file: 06-mlr-slides.html
    footer: "[🔗 sta101-f26.github.io](https://sta101-f26.github.io/)"
    theme: slides.scss
    transition: fade
    slide-number: true
    logo: images/logo.png
    toc: false
  html: 
    code-link: true
---


# Recap: simple linear regression

## From last time...penguins

```{r}
#| include: false
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(scatterplot3d)
library(ggthemes)
library(ggbeeswarm)
```

::: question
A researcher wants to see how body mass varies with flipper length.
:::

. . .

-   outcome: body mass in grams (numerical)

-   predictor: flipper length in mm (numerical)

. . .

Flipper length is easier to measure, so more plausible you would predict body mass based on that and not the other way around.

## How do we concisely summarize the association between two variables? 

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point()
```

## Line of best fit!

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Concise numerical summary: correlation

```{r}
penguins |>
  summarize(
    r = cor(flipper_length_mm, body_mass_g, use = "complete.obs")
  )
```

Measures the strength and direction of the linear association between two numerical variables. Strong and positive in this case.

## Model fit 

```{r}
bm_fl_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm, data = penguins)

tidy(bm_fl_fit)
```

$$
\widehat{body~mass}=-5780.83+49.68\times flipper~length
$$

## $R^2$ (coefficient of determination)

The fraction of the variation in the response explained by the model. A number between 0 (bad) and 1 (good) that measures goodness-of-fit:

:::::: columns
:::: {.column width="50%"}
```{r}
bm_fl_fit |>
  glance() |>
  select(r.squared)
```
:::

:::: {.column width="50%"}
```{r}
penguins |>
  summarize(
    r = cor(flipper_length_mm, body_mass_g, use = "complete.obs"),
    R2 = r^2
  )
```
:::

::::::

## Interpretation

$$
\widehat{body~mass}=-5780.83+49.68\times flipper~length
$$

-   Intercept: Penguins with 0 flipper length are expected, on average, to weigh -5,781 grams (makes no sense!);
-   Slopes: For each additional millimeter of a penguin's flipper length, the weight of the penguin is expected to be higher, on average, by 49.7 grams.

## Prediction 

```{r}
#| echo: false

new_penguin <- tibble(flipper_length_mm = 205)

preds <- predict(bm_fl_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point() + 
  geom_smooth(method = "lm")   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "#FE5D26"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "#FE5D26"
  )
```

## Prediction

```{r}
#| label: predict-210
new_penguin <- tibble(flipper_length_mm = 205)

predict(bm_fl_fit, new_data = new_penguin)
```

$$
\widehat{body~mass}=-5780.83+49.68\times 205\approx 4404.71
$$

. . .

We predict that a penguin whose flipper is 205 mm long will weigh 4404.71 grams on average.

# Linear regression with one categorical predictor

## New question

::: question
A different researcher wants to look at body weight of penguins based on the island they were recorded on. How are the variables involved in this analysis different?
:::

. . .

-   outcome: body mass in grams (numerical)

-   predictor: island (categorical with three levels)

. . .

```{r}
levels(penguins$island)
```

. . .

How should we visualize?


## Scatterplot: possible, but not so good 

::: task
Visualize the relationship between body weight and island of penguins. Also calculate the average body weight per island.
:::

```{r}
ggplot(penguins, aes(x = island, y = body_mass_g)) + 
  geom_point()
```

## Boxplot {.smaller}

::: question
Visualize the relationship between body weight and island of penguins.
Also calculate the average body weight per island.
:::

```{r}
#| output-location: column
#| warning: false
ggplot(
  penguins,
  aes(x = island, y = body_mass_g)
) +
  geom_boxplot() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body mass by island"
  )
```

```{r}
#| output-location: column
penguins |>
  group_by(island) |>
  summarize(
    bm_mean = mean(body_mass_g, na.rm = TRUE)
  )
```



## Density plot {.smaller}

```{r}
#| output-location: column
#| warning: false
ggplot(
  penguins,
  aes(x = body_mass_g, color = island, fill = island)
) +
  geom_density(alpha = 0.5) +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body mass by island"
  )
```


## Violin plot {.smaller}

```{r}
#| output-location: column
#| warning: false
ggplot(
  penguins,
  aes(x = island, y = body_mass_g)
) +
  geom_violin() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body mass by island"
  )
```

## Multiple geoms {.smaller}

```{r}
#| output-location: column
#| warning: false
#| fig-asp: 0.6
ggplot(
  penguins,
  aes(x = island, y = body_mass_g, color = island)
) +
  geom_boxplot() +
  geom_beeswarm(size = 2, alpha = 0.5) +
  scale_color_colorblind() +
  labs(
    x = "Island",
    y = "Body mass (g)",
    title = "Body mass by island"
  ) +
  theme(legend.position = "none")
```



## Model - fit {.medium}

::: task
Fit a linear regression model predicting body weight from island and display the results. Why is Biscoe not on the output?
:::

```{r}
bm_island_fit <- linear_reg() |>
  fit(body_mass_g ~ island, data = penguins)

tidy(bm_island_fit)
```

. . . 

Huh?

## Dummy variables {.medium}

```{r}
dummy_penguins <- penguins |>
  select(body_mass_g, island) |>
  arrange(body_mass_g) |>
  mutate(
    islandDream = if_else(island == "Dream", 1, 0),
    islandTorgersen = if_else(island == "Torgersen", 1, 0),
  )

dummy_penguins
```

## Dummy variables {.scrollable}

```{r}
linear_reg() |>
  fit(
    body_mass_g ~ islandDream + islandTorgersen, 
    dummy_penguins
  ) |>
  tidy()
```

. . .

Exactly the same as before!

## Model - interpret {.small}

$$
\widehat{body~mass} = 4716 - 1003 \times islandDream - 1010 \times islandTorgersen
$$

-   Intercept: [Penguins from Biscoe island are expected to weigh, on average, 4,716 grams.]{.fragment}

-   Slope - islandDream: [Penguins from Dream island are expected to weigh, on average, 1,003 grams less than those from Biscoe island.]{.fragment}

-   Slope - islandTorgersen: [Penguins from Torgersen island are expected to weigh, on average, 1,010 grams less than those from Biscoe island.]{.fragment}

## Model - predict {.medium .scrollable}

::: task
What is the predicted body weight of a penguin on Biscoe island? What are the estimated body weights of penguins on Dream and Torgersen islands? Where have we seen these values before?
:::

```{r}
new_penguins = tibble(
  island = c("Biscoe", "Dream", "Torgersen")
)

predict(bm_island_fit, new_data = new_penguins)
```

## Model - predict {.small .scrollable}

::: task
Calculate the predicted body weights of penguins on Biscoe, Dream, and Torgersen islands *by hand*.
:::

$$
\widehat{body~mass} = 4716 - 1003 \times islandDream - 1010 \times islandTorgersen
$$

. . .

-   Biscoe: $\widehat{body~mass} = 4716 - 1003 \times 0 - 1010 \times 0 = 4716$

. . .

-   Dream: $\widehat{body~mass} = 4716 - 1003 \times 1 - 1010 \times 0 = 3713$

. . .

-   Torgersen: $\widehat{body~mass} = 4716 - 1003 \times 0 - 1010 \times 1 = 3706$

## Look familiar?

```{r}
penguins |>
  group_by(island) |>
  summarize(
    bm_mean = mean(body_mass_g, na.rm = TRUE)
  )
```

## Models with categorical predictors

::: incremental
-   When the categorical predictor has many levels, they're encoded as **dummy variables**.

-   The first level of the categorical variable is the baseline level. In a model with one categorical predictor, the intercept is the predicted value of the outcome for the baseline level (x = 0).

-   Each slope coefficient describes the difference between the predicted value of the outcome for that level of the categorical variable compared to the baseline level.
:::

## In general {.small .scrollable}

Predicting continuous outcome $Y$ using one categorical predictor $X$ with multiple levels 0, 1, 2, ..., $L$. Create dummy variables for every level *except* the base level:

. . .

$$
cat_l=\begin{cases}
1 & X=l\\
0 & \text{else}
\end{cases}
$$

. . .

Then fit a regression with *multiple* dummy predictors:

$$
\hat{Y} = b_0 + b_1 \times cat_1 + b_2 \times cat_2 \ldots + b_L \times cat_L
$$

::: incremental
-   $b_0$ : the model prediction for a member of the base level;


-   $b_1$: how does the prediction change when we move from the base level to level 1?

- $b_2$: how does the prediction change when we move from the base level to level 2?

- etc
:::

. . .

::: callout-important
## We're not animals. We have technology!

The computer handles *all* of this for you, but you need to understand the details so you code and interpret it correctly.

:::

## Changing the baseline level

-   By default, R uses the first *level* of a categorical variable as the baseline level. This is often the first alphabetically, but make sure you check!

-   We can change the baseline level by reordering the levels of the categorical variable.

## Changing the baseline level


What is the baseline right now?

```{r}
levels(penguins$island)
```

. . .

Change it:

```{r}
penguins <- penguins |>
  mutate(
    island = fct_relevel(island, "Dream")
  )
```

. . .

What is the baseline now?

```{r}
levels(penguins$island)
```


## Another way to tell

```{r}
linear_reg() |>
  fit(body_mass_g ~ island, data = penguins) |>
  tidy()
```

Whichever level *is not* listed is treated as the baseline.


# Linear regression with one numerical and one categorical predictor

## Three variables on one 2D plot

```{r}
#| include: false 

penguins <- palmerpenguins::penguins
```

```{r}
ggplot(penguins, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point()
```

## How do we predict using more than one predictor? {.small}

Both of these models use `flipper_length_mm` *and* `island` to predict `body_mass_g`:


```{r}
#| label: additive-interaction-viz
#| layout-ncol: 2
#| echo: false
#| fig-asp: 1

# Plot A
bm_fl_island_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + island, data = penguins)
bm_fl_island_aug <- augment(bm_fl_island_fit, new_data = penguins)
ggplot(
  bm_fl_island_aug, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm") +
  labs(title = "Plot A - Additive model") +
  theme(legend.position = "bottom")

# Plot B
ggplot(
  penguins, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Plot B - Interaction model") +
  theme(legend.position = "bottom")
```

## The additive model: parallel lines, one for each island {.small .scrollable}

```{r}
#| label: additive-viz
#| echo: false


# Plot A
bm_fl_island_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + island, data = penguins)
bm_fl_island_aug <- augment(bm_fl_island_fit, new_data = penguins)
ggplot(
  bm_fl_island_aug, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm") +
  labs(title = "Additive model") +
  theme(legend.position = "bottom")


```

## The additive model: parallel lines, one for each island {.small .scrollable}

```{r}
bm_fl_island_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + island, data = penguins)

tidy(bm_fl_island_fit)
```

$$
\begin{aligned}
\widehat{body~mass} = -4625 &+ 44.5 \times flipper~length \\
&- 262 \times Dream \\
&- 185 \times Torgersen
\end{aligned}
$$

## Where do the three lines come from? {.scrollable}

$$
\begin{aligned}
\widehat{body~mass} = -4625 &+ 44.5 \times flipper~length \\
&- 262 \times Dream \\
&- 185 \times Torgersen
\end{aligned}
$$

. . .

If penguin is from Biscoe, Dream = 0 and Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{body~mass} = -4625 &+ 44.5 \times flipper~length
\end{aligned}
$$

. . .

If penguin is from Dream, Dream = 1 and Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{body~mass} = -4887 &+ 44.5 \times flipper~length
\end{aligned}
$$

. . .

If penguin is from Torgersen, Dream = 0 and Torgersen = 1:

. . .

$$
\begin{aligned}
\widehat{body~mass} = -4810 &+ 44.5 \times flipper~length
\end{aligned}
$$

. . .

Either way, same slope, so the lines are parallel.

## The interaction model: different lines for each island {.medium .scrollable}

```{r}
#| label: interaction-viz
#| code-fold: true

ggplot(
  penguins, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Interaction model") +
  theme(legend.position = "bottom")
```


## The interaction model: different lines for each island {.medium .scrollable}

```{r}
bm_fl_island_int_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm * island, data = penguins)

tidy(bm_fl_island_int_fit) |> select(term, estimate)
```

. . .

$$
\begin{aligned}
\widehat{body~mass} = -5464 &+ 48.5 \times flipper~length \\
&+ 3551 \times Dream \\
&+ 3218 \times Torgersen \\
&- 19.4 \times flipper~length*Dream \\
&- 17.4 \times flipper~length*Torgersen
\end{aligned}
$$

## Where do the three lines come from? {.medium .scrollable}

$$
\begin{aligned}
\small\widehat{body~mass} = -5464 &+ 48.5 \times flipper~length \\
&+ 3551 \times Dream \\
&+ 3218 \times Torgersen \\
&- 19.4 \times flipper~length*Dream \\
&- 17.4 \times flipper~length*Torgersen
\end{aligned}
$$

. . .

If penguin is from Biscoe, Dream = 0 and Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{body~mass} = -5464 &+ 48.5 \times flipper~length
\end{aligned}
$$

. . .

If penguin is from Dream, Dream = 1 and Torgersen = 0:

. . .

$$
\begin{aligned}
\widehat{body~mass} &= (-5464 + 3551) + (48.5-19.4) \times flipper~length\\
&=-1913+29.1\times flipper~length.
\end{aligned}
$$

## Prediction on Biscoe island

```{r}
#| echo: false

new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Biscoe"
  )

preds <- predict(bm_fl_island_int_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = island)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "#FE5D26"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "#FE5D26"
  )
```

## Prediction on Biscoe island

```{r}
new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Biscoe"
)

predict(bm_fl_island_int_fit, new_data = new_penguin)
```

$$
\widehat{body~mass} = -5464 + 48.5 \times 205.
$$

## Prediction on Dream island

```{r}
#| echo: false

new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Dream"
  )

preds <- predict(bm_fl_island_int_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = island)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "forestgreen"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "forestgreen"
  )
```

## Prediction on Dream island

```{r}
new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Dream"
)

predict(bm_fl_island_int_fit, new_data = new_penguin)
```

$$
\widehat{body~mass} = (-5464 + 3551) + (48.5 - 19.4) \times 205.
$$


## Prediction on Torgersen island

```{r}
#| echo: false

new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Torgersen"
  )

preds <- predict(bm_fl_island_int_fit, new_data = new_penguin)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = island)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)   +
  annotate(
    "segment",
    x = new_penguin$flipper_length_mm[1], xend = new_penguin$flipper_length_mm[1], y = -Inf, yend = preds$.pred[1],
    color = "blue"
  ) +
  annotate(
    "segment",
    x = -Inf, xend = new_penguin$flipper_length_mm[1], y = preds$.pred[1], yend = preds$.pred[1],
    color = "blue"
  )
```

## Prediction on Torgersen island

```{r}
new_penguin <- tibble(
  flipper_length_mm = 205,
  island = "Torgersen"
)

predict(bm_fl_island_int_fit, new_data = new_penguin)
```

$$
\widehat{body~mass} = (-5464 + 3218) + (48.5 - 17.4) \times 205.
$$


# Linear regression with many numerical predictors


## Multiple numerical predictors {.scrollable}

```{r}
bm_fl_bl_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + bill_length_mm, data = penguins)

tidy(bm_fl_bl_fit)
```

. . .

$$
\small\widehat{body~mass}=-5736+48.1\times flipper~length+6\times bill~length
$$


## Interpretation {.scrollable .medium}

$$
\small\widehat{body~mass}=-5736+48.1\times flipper~length+6\times bill~length
$$

. . .

**Interpretations**: 

::: incremental
- We predict that the body mass of a penguin with *zero* flipper length and *zero* bill length will be -5736 grams, on average (**makes no sense**);
- Holding all other variables constant, for every additional millimeter in flipper length, we expect the body mass of penguins to be higher, on average, by 48.1 grams.
- Holding all other variables constant, for every additional millimeter in bill length, we expect the body mass of penguins to be higher, on average, by 6 grams.
:::

## Prediction

```{r}
new_penguin <- tibble(
  flipper_length_mm = 200,
  bill_length_mm = 45
)

predict(bm_fl_bl_fit, new_data = new_penguin)
```

$$
\widehat{body~mass}=-5736+48.1\times 200+6\times 45
$$

## Picture? It's not pretty...

2 predictors + 1 response = 3 dimensions. Ick!

```{r}
#| echo: false
scatterplot3d(
  penguins[,c("bill_length_mm", "flipper_length_mm", "body_mass_g")],
  pch = 19, 
  color = "blue",
  xlab = "Bill length (mm)", 
  ylab = "Flipper length (mm)", 
  zlab = "Body mass (grams)"
) 
```

## Picture? It's not pretty...

Instead of a line of best fit, it's a *plane* of best fit. Double ick!

```{r}
#| echo: false
plot3vars <- scatterplot3d(
  penguins[,c("bill_length_mm", "flipper_length_mm", "body_mass_g")],
  pch = 19, 
  color = "blue",
  xlab = "Bill length (mm)", 
  ylab = "Flipper length (mm)", 
  zlab = "Body mass (grams)"
) 

model <- lm(body_mass_g ~ bill_length_mm + flipper_length_mm, data = penguins)
plot3vars$plane3d(model, col = "darkred")
```

# Multiple linear regression in general

## Population model {.medium}

**Multiple linear regression** captures the relationship between a numerical outcome $Y$ and *many* numerical predictors $X_1$, $X_2$, ..., $X_p$:

$$\Large{Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + ...+ \beta_p X_p+\epsilon}$$

::: incremental
-   $\beta_0$: True intercept of the relationship between the $X_j$ and $Y$
-   $\beta_j$: True slope of the relationship between $X_j$ and $Y$
-   $\epsilon$: Error (residual)
:::

. . .

The model with the greek letters and the error term is the "true," idealized, *population* relationship that we could access if we had infinite amounts of perfect data. But we don't, so we have to settle for...

## Estimated model {.medium}

$$\Large{\hat{Y} = b_0 + b_1 X_1 + b_2 X_2 + ... + b_pX_p}$$

-   $b_j$: Estimated slope of the relationship between $X_j$ and $Y$;
-   $b_0$: your prediction for $Y$ if all the predictors equal zero;
-   No error term!

. . .

This is your *best guess* at the true regression function based on the noisy, meager, imperfect data you actually have access to. We still compute the $b_j$ using the principle of least squares: pick the estimates that make the sum of squared residuals as small as possible.

## Next time

Today we saw multiple models that are all attempting to do the same thing: predict body mass.

- Which one is "best"?
- What does "best" even mean?

# The anatomy of the additive model

```{r}
#| label: anatomy-setup
#| echo: false

set.seed(1234567)
my_colors <- c("red", "blue", "forestgreen")
subsamps <- c(10, 20, 35)
b0 <- 12 + c(2, 6, -3)
b1 <- c(1, 5, -0.2)

y <- c()
x <- c()
cl <- c()

for(i in 1:length(my_colors)){
  x_sub <- rnorm(subsamps[i], 4, 1)
  y <- c(y, b0[i] + b1[i] * x_sub + rnorm(subsamps[i], 0, 1))
  x <- c(x, x_sub)
  cl <- c(cl, rep(my_colors[i], subsamps[i]))
}

df <- tibble(
  x = x, y = y, color = fct_relevel(cl, my_colors)
)
df <- df[sample(1:nrow(df)), ]

add_slope_wedge <- function(intercept, slope, x0,
                            run = NULL,       # horizontal length of wedge
                            cex = 1,          # text size
                            col = "black",
                            lwd = 2,
                            label = TRUE) {

  # If run not supplied, choose something that looks reasonable
  if (is.null(run)) {
    usr <- par("usr")              # current plot limits
    run <- 0.08 * (usr[2] - usr[1])
  }
  
  # Point on the line where wedge attaches
  y0 <- intercept + slope * x0
  
  # End of horizontal "run"
  x1 <- x0 + run
  y1 <- y0
  
  # Vertical "rise"
  rise <- slope * run
  x2 <- x1
  y2 <- y1 + rise
  
  # Draw the triangle edges
  segments(x0, y0, x1, y1, col = col, lwd = lwd)  # run
  segments(x1, y1, x2, y2, col = col, lwd = lwd)  # rise
  segments(x0, y0, x2, y2, col = col, lwd = lwd)  # hypotenuse (optional but nice)
  
  # Optional labels
  if (label) {
    text((x0 + x1)/2, y0, "run", pos = 1, cex = cex, col = col)
    text(x1, (y1 + y2)/2, "rise", pos = 4, cex = cex, col = col)
  }
}

my_fit <- linear_reg() |>
  fit(y ~ x + color, df) |>
  tidy()

cepts <- c(my_fit$estimate[1], my_fit$estimate[1] + my_fit$estimate[3], my_fit$estimate[1] + my_fit$estimate[4])
slope <- my_fit$estimate[2]
```

## Toy dataset

:::: columns

::: column
Consider this nonsense dataset with three variables:

- `x`: numerical predictor;
- `y`: numerical response;
- `color`: categorical predictor with three levels (red, green, blue)
:::

:::column
```{r}
df
```
:::


:::: 

## Additive model


:::: columns

::: column
Fit the *additive* model:

```{r}
#| eval: false
linear_reg() |>
  fit(y ~ x + color, df) |>
  tidy()
```
:::

:::column
You get a table with the coefficient estimates:

| `term` | `estimate` | 
|---------|:-----|
| `(Intercept)`      | $b_0$   |   
| `x`     | $b_1$  |  
| `colorblue`       | $b_2$   |   
| `colorforestgreen`       | $b_3$    | 
:::


:::: 

So `red` is apparently the *baseline* level.

## The raw data

```{r}
#| echo: false
plot(x, y, col = cl, pch = 19, xaxt = "n", yaxt = "n", bty = "n", xlim = c(-2, 1.4 * max(x)), ylim = c(-5, max(y)), cex = 1)
abline(v = 0, lwd = 2)
abline(h = 0, lwd = 2)
```

## Model fit: parallel lines, one for each group

```{r}
#| echo: false

plot(x, y, col = cl, pch = 19, xaxt = "n", yaxt = "n", bty = "n", xlim = c(-2, 1.4 * max(x)), ylim = c(-5, max(y)), cex = 1)
abline(v = 0, lwd = 2)
abline(h = 0, lwd = 2)
abline(a = cepts[1], b = slope, col = my_colors[1], lwd = 2)
abline(a = cepts[2], b = slope, col = my_colors[2], lwd = 2)
abline(a = cepts[3], b = slope, col = my_colors[3], lwd = 2)

```

## The gory details

```{r}
#| echo: false

plot(x, y, col = cl, pch = 19, xaxt = "n", yaxt = "n", bty = "n", xlim = c(-2, 1.4 * max(x)), ylim = c(-5, max(y)), cex = 1)
add_slope_wedge(cepts[1], slope, 7, label = FALSE)
add_slope_wedge(cepts[2], slope, 7, label = FALSE)
add_slope_wedge(cepts[3], slope, 7, label = FALSE)
abline(v = 0, lwd = 2)
abline(h = 0, lwd = 2)
abline(a = cepts[1], b = slope, col = my_colors[1], lwd = 2)
abline(a = cepts[2], b = slope, col = my_colors[2], lwd = 2)
abline(a = cepts[3], b = slope, col = my_colors[3], lwd = 2)
points(numeric(length(cepts)), 
       cepts, 
       cex = 2, pch = 19)
arrows(-1, 0.95 * cepts[2] + slope * (-1), -1, 1.1 * cepts[1] + slope * (-1), lwd = 2, code = 3)
arrows(-1, 0.95 * cepts[1] + slope * (-1), -1, 1.1 * cepts[3] + slope * (-1), lwd = 2, code = 3)
#arrows(-3, 0.95 * cepts[2] + slope * (-3), -3, 1.1 * cepts[3] + slope * (-3), lwd = 2, code = 3) # the big arrow


#text(0.5, 1.5 + cepts[2] + slope * 0.5, "IV")

#text(-4, 0.5*cepts[1] + 0.5*cepts[2], "I")
text(-2, 0.5*cepts[1] + 0.5*cepts[2], expression(b[2]))
text(-2, 0.5*cepts[1] + 0.5*cepts[3], expression("|"~b[3]~"|"))

text(0.5, 2 + cepts[2] + slope * 0.5, expression(b[0]~" + "~b[2]))
text(0.5, 2 + cepts[1] + slope * 0.5, expression(b[0]))
text(0.5, 2 + cepts[3] + slope * 0.5, expression(b[0]~" + "~b[3]))

text(7.5, cepts[2] + slope * 7 - 2, expression(b[1]))
text(7.5, cepts[1] + slope * 7 - 2, expression(b[1]))
text(7.5, cepts[3] + slope * 7 - 2, expression(b[1]))
```

## The gory details

Best fit line for red:

$$
\hat{y}=b_0+b_1 x.
$$

Best fit line for blue:

$$
\hat{y}=(b_0 + b_2)+b_1 x.
$$

Best fit line for green:

$$
\hat{y}=(b_0 + b_3)+b_1 x.
$$

# The anatomy of the interaction model

```{r}
#| label: anatomy-setup-2
#| echo: false
#| fig-asp: 1

set.seed(1234567)
my_colors <- c("red", "blue", "forestgreen")
my_shapes <- my_colors#c("square", "circle", "triangle")
my_codes <- 15:17
#lookup <- setNames(my_codes, my_shapes)
#subsamps <- c(35, 20, 15)
b0 <- 10 + c(0, 40, 15)
b1 <- c(2, -2, 1)

y <- c()
x <- c()
cl <- c()

for(i in 1:length(my_shapes)){
  x_sub <- rnorm(subsamps[i], 4, 1)
  y <- c(y, b0[i] + b1[i] * x_sub + rnorm(subsamps[i], 0, 1.5))
  x <- c(x, x_sub)
  cl <- c(cl, rep(my_shapes[i], subsamps[i]))
}

df <- tibble(
  x = x, y = y, color = fct_relevel(cl, my_shapes)
)
df <- df[sample(1:nrow(df)), ]


add_slope_wedge <- function(intercept, slope, x0,
                            run = NULL,       # horizontal length of wedge
                            cex = 1,          # text size
                            col = "black",
                            lwd = 2,
                            label = TRUE) {

  # If run not supplied, choose something that looks reasonable
  if (is.null(run)) {
    usr <- par("usr")              # current plot limits
    run <- 0.08 * (usr[2] - usr[1])
  }
  
  # Point on the line where wedge attaches
  y0 <- intercept + slope * x0
  
  # End of horizontal "run"
  x1 <- x0 + run
  y1 <- y0
  
  # Vertical "rise"
  rise <- slope * run
  x2 <- x1
  y2 <- y1 + rise
  
  # Draw the triangle edges
  segments(x0, y0, x1, y1, col = col, lwd = lwd)  # run
  segments(x1, y1, x2, y2, col = col, lwd = lwd)  # rise
  segments(x0, y0, x2, y2, col = col, lwd = lwd)  # hypotenuse (optional but nice)
  
  # Optional labels
  if (label) {
    text((x0 + x1)/2, y0, "run", pos = 1, cex = cex, col = col)
    text(x1, (y1 + y2)/2, "rise", pos = 4, cex = cex, col = col)
  }
}

my_fit <- linear_reg() |>
  fit(y ~ x * color, df) |>
  tidy()

cepts <- c(my_fit$estimate[1], my_fit$estimate[1] + my_fit$estimate[3], my_fit$estimate[1] + my_fit$estimate[4])
slopes <- c(my_fit$estimate[2], my_fit$estimate[2] + my_fit$estimate[5], my_fit$estimate[2] + my_fit$estimate[6])
slope <- my_fit$estimate[2]
```

## Another toy dataset

:::: columns

::: column
Consider this nonsense dataset with three variables:

- `x`: numerical predictor;
- `y`: numerical response;
- `color`: categorical predictor with three levels (red, green, blue)
:::

:::column
```{r}
df
```
:::


:::: 

## Interaction model


:::: columns

::: column
Fit the *interaction* model:

```{r}
#| eval: false
linear_reg() |>
  fit(y ~ x * color, df) |>
  tidy()
```
:::

:::column
You get a table with the coefficient estimates:

| `term` | `estimate` | 
|---------|:-----|
| `(Intercept)`      | $b_0$   |   
| `x`     | $b_1$  |  
| `colorblue`       | $b_2$   |   
| `colorforestgreen`       | $b_3$    |  
| `x:colorblue`       | $b_4$    |   
| `x:colorforestgreen`       | $b_5$    |  
:::


:::: 

So `red` is apparently the *baseline* level.

## The raw data


```{r}
#| echo: false
par(mar = c(4, 4, 0.1, 0.1))

plot(x, y, col = cl, pch = 19, xaxt = "n", yaxt = "n", bty = "n", xlim = c(-5, 1.1 * max(x)), ylim = c(-6, 1.3 * max(y)), cex = 1.5)
abline(v = 0, lwd = 2)
abline(h = 0, lwd = 2)
```

## Model fit: a line for each group, but crossing


```{r}
#| echo: false
par(mar = c(4, 4, 0.1, 0.1))

plot(x, y, col = cl, pch = 19, xaxt = "n", yaxt = "n", bty = "n", xlim = c(-5, 1.1 * max(x)), ylim = c(-6, 1.3 * max(y)), cex = 1.5)
abline(v = 0, lwd = 2)
abline(h = 0, lwd = 2)
abline(a = cepts[1], b = slopes[1], lwd = 2, col = my_colors[1])
abline(a = cepts[2], b = slopes[2], lwd = 2, col = my_colors[2])
abline(a = cepts[3], b = slopes[3], lwd = 2, col = my_colors[3])
```

## The gory details


```{r}
#| echo: false
par(mar = c(4, 4, 0.1, 0.1))

#plot(x, y, pch = unname(lookup[cl]), xaxt = "n", yaxt = "n", bty = "n", xlim = c(-5, 1.1 * max(x)), ylim = c(-5, 1.3 * max(y)), cex = 1)
plot(x, y, col = cl, pch = 19, xaxt = "n", yaxt = "n", bty = "n", xlim = c(-5, 1.1 * max(x)), ylim = c(-6, 1.3 * max(y)), cex = 1.5)
add_slope_wedge(cepts[1], slopes[1], -5, label = FALSE)
add_slope_wedge(cepts[2], slopes[2], -5, label = FALSE)
add_slope_wedge(cepts[3], slopes[3], -5, label = FALSE)
abline(v = 0, lwd = 2)
abline(h = 0, lwd = 2)
abline(a = cepts[1], b = slopes[1], lwd = 2, col = my_colors[1])
abline(a = cepts[2], b = slopes[2], lwd = 2, col = my_colors[2])
abline(a = cepts[3], b = slopes[3], lwd = 2, col = my_colors[3])
points(numeric(length(cepts)), 
       cepts, 
       cex = 2, pch = 19)

segments(rep(-300, length(cepts)), cepts, numeric(length(cepts)), cepts, lty = 2)

arrows(-3, cepts[2], -3, cepts[1], lwd = 2, code = 3) # big arrow
arrows(-1, cepts[1], -1, 0.95 * cepts[3] + slopes[3] * (-1), lwd = 2, code = 3) # bottom arrow
#arrows(-1, 0.95 * cepts[2] + slopes[2] * (-1), -1, 1.1 * cepts[3] + slopes[3] * (-1), lwd = 2, code = 3)

text(-5, 3 + cepts[1] + slopes[1] * (-5), expression(b[1]))
text(-5, -3 + cepts[2] + slopes[2] * (-5), expression(b[1]~" + "~b[4]))
text(-5, -3 + cepts[3] + slopes[3] * (-5), expression(b[1]~" + "~b[5]))

text(-3.5, (cepts[2] + cepts[3]) / 2, expression(b[2]))

#text(-1.75, (cepts[2] + cepts[3]) / 2, paste0(q_no + 5))
text(-1.75, (cepts[1] + cepts[3]) / 2, expression(b[3]))

text(0.5, 2 + cepts[2] + slope * 0.5, expression(b[0]~" + "~b[2]))
text(0.5, 2 + cepts[3] + slope * 0.5, expression(b[0]~" + "~b[3]))
text(0.5, 2 + cepts[1] + slope * 0.5, expression(b[0]))
```


## The gory details

Best fit line for red:

$$
\hat{y}=b_0+b_1 x.
$$

Best fit line for blue:

$$
\hat{y}=(b_0 + b_2)+(b_1 + b_4)x.
$$

Best fit line for green:

$$
\hat{y}=(b_0 + b_3)+(b_1+b_5)x.
$$
