---
title: "Midterm Study Guide"
format: html
execute:
  echo: false
---

```{r}
#| label: load-packages
#| message: false
#| warning: false
#| echo: false
library(tidyverse)
library(tidymodels)
library(openintro)
library(scales)
library(palmerpenguins)
library(MASS)
library(conflicted)
conflicts_prefer(
  dplyr::select(),
  dplyr::filter(),
  palmerpenguins::penguins
)
theme_set(theme_minimal(base_size = 12))

```

Our midterm exam will take place during lecture on Thursday October 8. Here's what you should know about it:

- **What's the format?** The exam is a traditional pencil-and-paper exam. The questions will be a mix of multiple choice and free response;
- **How long will it take?** You have 75 minutes to complete the exam, but on average the exam is not *intended* to take the full 75 minutes;
- **What resources are allowed?** You are only allowed both sides of a single 8.5" x 11" sheet of notes created by you and you only. I do not care how you create the note sheet;
- **Will we be writing code on the exam?** No. However, you may be asked to analyze and interpret the *output* of some code, and so being able to *read* the code that generated the output could be helpful;
- **What skills will be emphasized?** JZ cares most about your visual intuition and conceptual understanding, so you should sweep through all of the lectures, find every *picture* we ever dissected, and make sure you understand every detail of it;
- **What topics will be covered?** Everything from Lecture 2 on 9/1 through Lecture 11 on 10/6 is fair game. So that's exploratory data analysis (visualization and summarization) and regression modeling (both linear and logistic);
- **Should I study the textbook, the lecture videos, or the tutorials?** Maybe if you have time, but you should not prioritize them. In approximate order of importance, you should prioritize studying this guide, the lecture notes, the solutions to the in-class demos (so-called AEs), the homeworks, and the labs. If you have more time to study after all of that, then the first thing I would suggest is sampling some of the odd-numbered exercises at the end of Ch. 1 - 10 in [IMS](https://openintrostat.github.io/ims/). The answers to these exercises are in the book, and they do not involve code: just concepts, interpretation, etc;
    - **Note**: if you do dip into the textbook, videos, or tutorials, and you encounter a topic that *never appeared* in any of the lectures, labs, or homeworks, then you are not responsible for it;
- **One more thing:** don’t waste your time studying the details of any specific applications or datasets we’ve seen (eg. penguins, Bechdel, gerrymandering, diamonds, etc). Real datasets will of course be discussed on the exam, but everything you need to know about them will be introduced from scratch within the exam as if you had never seen them before.

Below are a bunch of practice problems to get your wheels spinning. Solutions will be posted on Monday October 5.

::: callout-tip
## Try this

Attempt all of the problems below *without resources*. Make note of your pain points and blind spots along the way. That might give you an idea of the stuff you would benefit from if you had it on a cheat sheet.
:::


# Concepts and main ideas

```{r}
question <- 1
```

### Question `{r} question`


Can you answer each of these in one or two sentences?

i. If I have a column of numbers, what is the definition of the 38% quantile? 
    - Could you sketch a cartoon of what this looks like? 
    - Is the 38% quantile smaller than, bigger than, or equal to the median, or do you not have enough information to say?
i. What does the correlation coefficient between two numerical variables measure?
i. What is the difference between $y=\beta_0+\beta_1 x + \varepsilon$ and $\hat{y}=b_0+b_1 x$?
i. When we plot a line of best fit for a numerical predictor and numerical response, what does "best" mean?
i. Consider this bit of toy code: `linear_reg() |> fit(_BLANK_, df)`? What goes in the blank? What purpose does that argument serve, and how do you use it?
i. Under what circumstances is the intercept in a regression model a meaningful quantity?
i. If you run a "simple" linear regression model with a numerical response and a categorical predictor with three levels, what does the computer actually do to estimate that model?
i. Let $\hat{y}^{(\text{old})}$ be the prediction at $x$ of a simple linear regression, and let $\hat{y}^{(\text{new})}$ be the prediction at $x+1$ of the same regression. How are $\hat{y}^{(\text{old})}$ and $\hat{y}^{(\text{new})}$ related? In other words, how does the prediction of the simple linear regression model change when you change the input from $x$ to $x+1$?
i. What does the $R^2$ of a linear regression measure?
i. Why is adjusted $R^2$ a better model selection criterion than ordinary, unadjusted $R^2$?
i. Let $\hat{o}^{(\text{old})}$ be the estimated odds at $x$ from a simple logistic regression model, and let $\hat{o}^{(\text{new})}$ be the estimated odds at $x+1$ of the same regression. How are $\hat{o}^{(\text{old})}$ and $\hat{o}^{(\text{new})}$ related? In other words, how does the prediction of the simple logistic regression model change when you change the input from $x$ to $x+1$?
i. What is the main lesson of Anscombe's Quartet?

# Visual understanding

```{r}
question <- question + 1
```

### Question `{r} question`

Practice!

<https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>

```{r}
question <- question + 1
```

### Question `{r} question`

```{r}
#| include: false

set.seed(123)

# ------------------------------------------------------------
# Define 6 distinct 3x3 count tables (Group x Response)
# ------------------------------------------------------------

tables <- list(

  # 1: Strong gradient across groups, zero in G1-C
  "i" = tribble(
    ~Group, ~Yes, ~Maybe, ~No,
    "G1",    60, 30,  0,
    "G2",    30, 40, 30,
    "G3",    10, 30, 60
  ),

  # 2: Same general direction as 1 but very different totals + different shape
  "ii" = tribble(
    ~Group, ~Yes, ~Maybe, ~No,
    "G1",    30, 15,  5,
    "G2",    25, 25, 25,
    "G3",    10, 20, 70
  ),

  # 3: Near independence, very unequal totals
  "iii" = tribble(
    ~Group, ~Yes, ~Maybe, ~No,
    "G1",    20, 20, 20,
    "G2",    40, 40, 40,
    "G3",    10, 10, 10
  ),

  # 4: One dominant category per group (rotating dominance), includes zeros
  "iv" = tribble(
    ~Group, ~Yes, ~Maybe, ~No,
    "G1",    80, 10, 10,
    "G2",    10, 80, 10,
    "G3",    10, 10, 80
  ),

  # 5: Heavy imbalance in totals, subtle proportion differences
  "v" = tribble(
    ~Group, ~Yes, ~Maybe, ~No,
    "G1",    100, 50, 50,
    "G2",     40, 20, 40,
    "G3",     10,  5,  5
  ),

  # 6: Zeros scattered, nonlinear pattern
  "vi" = tribble(
    ~Group, ~Yes, ~Maybe, ~No,
    "G1",     0, 40, 60,
    "G2",    50,  0, 50,
    "G3",    60, 40,  0
  )
)

# ------------------------------------------------------------
# Convert to long format and expand to individual rows
# ------------------------------------------------------------

data_list <- lapply(names(tables), function(name) {
  tables[[name]] |>
    pivot_longer(cols = Yes:No,
                 names_to = "Response",
                 values_to = "n") |>
    uncount(n) |>
    mutate(Dataset = name)
})

df <- bind_rows(data_list)

mynames <- c("i", "ii", "iii", "iv", "v", "vi")

df$Dataset <- factor(df$Dataset, levels = mynames)
df$Group <- factor(df$Group, levels = c("G1","G2","G3"))
df$Response <- factor(df$Response, levels = c("Yes", "Maybe", "No"))

# ------------------------------------------------------------
# Create 3x2 numbered dodge plot panel
# ------------------------------------------------------------

dodge_plot <- ggplot(df, aes(x = Group, fill = Response)) +
  geom_bar(position = "dodge", color = "black") +
  facet_wrap(~Dataset, ncol = 3) +
  labs(title = "",
       x = "Group",
       y = "Count") +
  theme_bw() +
  theme(legend.position = "right") + 
  scale_fill_manual(
    values = c("Yes" = "gray30", "Maybe" = "gray80", "No" = "white")
  )

# ------------------------------------------------------------
# Create 3x2 lettered fill plot panel
# ------------------------------------------------------------

# Randomly relabel datasets A–F
letter_map <- tibble(
  Dataset = factor(mynames),
  Letter = factor(sample(letters[1:6]))
)

df_lettered <- df |>
  left_join(letter_map, by = "Dataset")

fill_plot <- ggplot(df_lettered, aes(x = Group, fill = Response)) +
  geom_bar(position = "fill", color = "black") +
  facet_wrap(~Letter, ncol = 3) +
  labs(title = "",
       x = "Group",
       y = "Proportion") +
  theme_bw() +
  theme(legend.position = "right") + 
  scale_fill_manual(
    values = c("Yes" = "gray30", "Maybe" = "gray80", "No" = "white")
  )

# ------------------------------------------------------------
# Print plots
# ------------------------------------------------------------


```

Match the "dodged" barchart with the corresponding stacked barchart. There is exactly one unique match for each.

```{r}
#| echo: false
#| fig-align: center
#| fig-asp: 0.53
dodge_plot
```


```{r}
#| echo: false
#| fig-align: center
#| fig-asp: 0.53
fill_plot
```

```{r}
question <- question + 1
```

### Question `{r} question`


The scatterplots below show the relationships between six pairs of `x` and `y` variables.

```{r}
#| label: simulate-data
#| echo: false
set.seed(8675309)

banana = function(A, B, C1, C2, N, keep = 10, init = 10)
{
  R=init*keep+N*keep
  x1=x2=0
  bimat=matrix(double(2*N),ncol=2)
  for (r in 1:R) {
    x1=rnorm(1,mean=(B*x2+C1)/(A*(x2^2)+1),sd=sqrt(1/(A*(x2^2)+1)))
    x2=rnorm(1,mean=(B*x2+C2)/(A*(x1^2)+1),sd=sqrt(1/(A*(x1^2)+1)))
    if (r>init*keep && r%%keep==0) {
        mkeep=r/keep; bimat[mkeep-init,]=c(x1,x2)
    }
  }

  return(bimat)
}

n <- 2500
p <- rbinom(n, 1, 0.5)
q <- rbinom(n, 1, 0.9)

z1 <- mvrnorm(n, numeric(2), 0.1 * cbind(c(1, 0.8), c(0.8, 1)))

z2 <- mvrnorm(n, c(-0.5, 0), 0.05 * cbind(c(1, 0), c(0, 2))) * q + 
      mvrnorm(n, c(0.5, 0), 0.05 * cbind(c(1, 0), c(0, 2))) * (1 - q)
z3 <- cbind(rnorm(n, mean = 0, sd = 0.35), 
            0.5 * rexp(n) - 1)
z4 <- mvrnorm(n, numeric(2), 0.1 * cbind(c(2, 0), c(0, 0.25)))
z5 <- mvrnorm(n, 0.5 + numeric(2), 0.01 * cbind(c(1, -0.5), c(-0.5, 1)))
z6 <- mvrnorm(n, c(-0.5, 0.5), 0.05 * cbind(c(1, 0), c(0, 1))) * p + 
      mvrnorm(n, c(0.5, -0.5), 0.05 * cbind(c(1, 0), c(0, 1))) * (1 - p)

plot_labels <- mynames

z <- tibble(
  plot_no = c(rep(plot_labels[1], n), 
              rep(plot_labels[2], n), 
              rep(plot_labels[3], n), 
              rep(plot_labels[4], n), 
              rep(plot_labels[5], n), 
              rep(plot_labels[6], n)),
  x = c(z1[,1], z2[,1], z3[,1], z4[,1], z5[,1], z6[,1]),
  y = c(z1[,2], z2[,2], z3[,2], z4[,2], z5[,2], z6[,2])
) |>
  mutate(plot_no = fct_relevel(plot_no, plot_labels))
```

```{r}
#| label: simulated-data-scatterplots
#| echo: false
#| warning: false
#| fig-asp: 0.66
#| fig-width: 8
ggplot(z, aes(x = x, y = y)) +
  geom_point(size = 0.25) +
  facet_wrap(~plot_no) +
  scale_x_continuous(limits = c(-1, 1)) +
  scale_y_continuous(limits = c(-1, 1)) +
  coord_fixed()
```

Match each scatterplot to the pair of boxplots that displays the distributions of `x` and `y` separately. There is exactly one unique match for each.

```{r}
#| label: simulated-data-histograms
#| echo: false
#| warning: false
#| fig-asp: 1.25
#| fig-width: 8
z |>
  pivot_longer(
    cols = c(x, y),
    names_to = "variable",
    values_to = "value"
  ) |>
  mutate(
    plot_no = case_when(
      plot_no == plot_labels[1]   ~ "Boxplot pair e",
      plot_no == plot_labels[2]  ~ "Boxplot pair d",
      plot_no == plot_labels[3] ~ "Boxplot pair b",
      plot_no == plot_labels[4]  ~ "Boxplot pair a",
      plot_no == plot_labels[5]  ~ "Boxplot pair c",
      plot_no == plot_labels[6]  ~ "Boxplot pair f",
    )
  ) |>
  ggplot(aes(x = value)) +
  geom_boxplot() +
  facet_grid(
    plot_no ~ variable, 
    switch = "y"
  ) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(x = NULL, y = NULL) +
  theme(
    strip.placement = "outside",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

::: callout-tip
## Hint
Notice that the range of the axes is the same in every picture.
:::


```{r}
question <- question + 1
```

### Question `{r} question`

In what follows, let $y$ be a binary response, let $x$ be a numerical predictor, and let $z$ be a binary predictor. Match the S-curve to the model equation.

```{r}
#| fig-asp: 1
#| fig-width: 8
#| warning: false

set.seed(1234)

par(mfrow = c(3, 2))

n <- 100
x <- seq(-10, 10, length.out = n)
z <- rbinom(n, 1, 0.5)

# normal

b0 <- 0
b1 <- 0.5
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs3 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs3[1]
b1 <- coeffs3[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot i")

# additive

b0 <- -3
b1 <- -1
b2 <- 2
y <- rbinom(n, 1, prob = exp(b0 + b1 * x + b2 * z) / (1 + exp(b0 + b1 * x + b2 * z)))

coeffs4 <- coefficients(glm(y ~ x + z, binomial(link = "logit")))
b0 <- coeffs4[1]
b1 <- coeffs4[2]
b2 <- coeffs4[3]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot ii")
curve(exp(b0 + b1 * x + b2) / (1 + exp(b0 + b1 * x + b2)), from = min(x), to = max(x), n = 1000, col = "blue", bty = "n", ylab = "", ylim = c(0, 1), add = TRUE)

# flipped

b0 <- 0
b1 <- -0.5
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs5 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs5[1]
b1 <- coeffs5[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot iii")

# shifted

b0 <- -5
b1 <- 1
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs6 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs6[1]
b1 <- coeffs6[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot iv")

# scrunched

b0 <- 20
b1 <- 4
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs7 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs7[1]
b1 <- coeffs7[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot v")


# stretched

b0 <- 0.5
b1 <- 0.25
y <- rbinom(n, 1, prob = exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)))

coeffs8 <- coefficients(glm(y ~ x, binomial(link = "logit")))
b0 <- coeffs8[1]
b1 <- coeffs8[2]

curve(exp(b0 + b1 * x) / (1 + exp(b0 + b1 * x)), from = min(x), to = max(x), n = 1000, col = "red", bty = "n", ylab = "Prob(y = 1 | x)", xlab = "x", ylim = c(0, 1), main = "Plot vi")

```

a. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 22.9 + 4.54 \times x$
a. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 0.118 + 0.246 \times x$
a. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 0.798 + 0.486 \times x$
a. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = -6.035 + 1.194 \times x$
a. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = -2.26 - 0.99 \times x+1.26\times z$
a. $\log\left(\frac{\hat{p}}{1-\hat{p}}\right) = 0.079  -0.373 \times x$

```{r}
question <- question + 1
```

### Question `{r} question`

```{r}
#| echo: false
email <- email |> relocate(spam, exclaim_mess)
```


Recall the dataset on spam emails:

```{r}
#| echo: true
email
```

Here is a regression model trained on all of these data:

```{r}
#| echo: true
#| warning: false
#| message: false
log_fit <- logistic_reg() |>
  fit(spam ~ ., data = email)
```

The *in-sample* predictions of the model are here:

```{r}
#| echo: true
email_aug <- augment(log_fit, email)
email_aug
```

To see how well the model did, we can compute the classification error rates for different thresholds and summarize the results with the ROC curve:

```{r}
#| echo: true
email_roc <- roc_curve(email_aug, truth = spam, .pred_1, event_level = "second")
```


```{r}
#| echo: false

ggplot(email_roc, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  annotate("point", x = c(0.0, 0.0819, 0.229, 1.0), 
                    y = c(0.0, 0.531, 0.872, 1.0), 
                    size = 5) + 
  annotate("text", x = c(0.0 + 0.1, 0.2, 0.35, 1.0), 
                   y = c(0.0, 0.5, 0.872, 1.0 - 0.1), 
                   label = c("i", "ii", "iii", "iv"), 
                   size = 5) +
  geom_abline(lty = 3) +
  coord_equal() + 
  labs(
    x = "1 - specificity",
    y = "sensitivity"
  )
```


Match the point on the ROC curve to the threshold that was used to compute it. 


a. 

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 0.25, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```

b. 

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 0.0, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```

c. 

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 1.00, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```

d.

```{r}
#| echo: true
email_aug |>
  mutate(.pred_class = if_else(.pred_1 <= 0.1, 0, 1)) |>
  count(spam, .pred_class) |> group_by(spam) |> mutate(prop = n / sum(n))
```


# STAWANOWA doctor

What's wrong with the interpretations below? Diagnose the problem and fix it if possible.

```{r}
q_no <- question + 1
```

## Question `{r} q_no`

Some bozo ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(bill_length_mm ~ flipper_length_mm, penguins) |>
  tidy()
```

> If the flipper length of a penguin increases by 1mm, then the bill length will increase by 0.25mm.

```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some schmuck ran this regression and wrote the following interpretation:

```{r}
#| echo: true
logistic_reg() |>
  fit(spam ~ num_char, email) |>
  tidy()
```

> If the number of characters in an email increases by 1000, then the estimated probability of that email being spam is lower by -0.062, on average.

FYI: from the documentation (`?email`), `num_char` is "The number of characters in the email, in thousands."

```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some putz ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + bill_length_mm, penguins) |> 
  tidy()
```

> If flipper length and bill length both increase by 1mm, then we predict that body mass is lower by 48.14 grams, on average.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some blockhead ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm * species, penguins) |>
  tidy()
```

> If the flipper length of a Chinstrap penguin increases by 1mm, then we predict that the body mass is higher by 1.74 grams on average.

```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some nitwit ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(hp ~ mpg, mtcars) |>
  glance()
```

> About 60% of the observed variation in miles/gallon is explained by the model.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some chooch ran this regression and wrote the following interpretation:

```{r}
#| echo: true
logistic_reg() |>
  fit(spam ~ exclaim_mess + viagra + dollar, email) |>
  tidy()
```

> If an email contains no exclamation marks, no dollar signs, and no mentions of viagra, then the model predicts that the probability of it being spam is $e^{-2.218}\approx 0.10886$.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some nincompoop ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(bill_length_mm ~ bill_depth_mm + sex, penguins) |>
  tidy()
```

> If a male penguin has a bill depth of 0mm (sad!), then we expect its bill length to be 60.99mm on average.


```{r}
q_no <- q_no + 1
```

## Question `{r} q_no`

Some silly goose ran this regression and wrote the following interpretation:

```{r}
#| echo: true
linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + bill_length_mm, penguins) |>
  glance()
```

> The correlation between body mass, flipper length, and bill length is about $\sqrt{0.76}\approx 0.87$.



# Data analysis

## Law & Order

You've heard of the tidyverse, now let's visit the Law & Order-verse. Doink doink![^1]

[^1]: "Doink doink" is the scene and episode introductory sound on the Law & Order series. If you've never heard it, you're not at any disadvantage for the exam. If you've ever heard it, good luck getting it out of your head!

Law & Order is a police procedural and legal drama television series that has been running since the 1990s. The Law & Order franchise includes a number of series such as Law & Order, Law & Order: SVU, Law & Order: Criminal Intent, etc.

You will work with data on average ratings for each season of three series from the Law & Order-verse – a subset of the data from the previous questions. Below is a peek at the first ten rows of the Law & Order data.

```{r}
#| echo: false
#| message: false
tv <- read_csv("data/tv.csv")
law_and_order <- tv |>
  filter(str_detect(title, "Law & Order"))
```

The plot below shows the distributions of average ratings of various Law & Order series across seasons.

```{r}
#| echo: false
#| fig-asp: 0.35
#| fig-width: 8

law_and_order |>
  mutate(title = str_replace(title, "Law & Order: ", "Law & Order:\n")) |>
  ggplot(aes(x = av_rating, y = fct_rev(title))) +
  geom_boxplot(alpha = 0.5, show.legend = FALSE, outlier.size = 2) +
  labs(x = "Average rating", y = "Series title")
```

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Based on the information from the side-by-side box plots, fill in the legend of the plot below with Law & Order series titles.

```{r}
#| echo: false
#| fig-asp: 0.35
#| fig-width: 8
#| warning: false

ggplot(law_and_order, aes(x = season, y = av_rating, shape = title)) +
  geom_line(linewidth = 0.3) +
  geom_point(size = 3) +
  scale_shape_manual(
    values = c("Law & Order: SVU" = "triangle", 
               "Law & Order: Criminal Intent" = "diamond", 
               "Law & Order: True Crime" = "square", 
               "Law & Order" = "asterisk", 
               "Law & Order: LA" = "circle"), 
    labels = c("__________________________________",
               "__________________________________",
               "__________________________________",
               "__________________________________",
               "__________________________________")
  ) +
  labs(
    x = "Season",
    y = "Average rating",
    shape = "Series title"
  ) +
  theme(legend.key.height = unit(1.25, "cm"))
```

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

The following code calculates the standard deviations of average season ratings of the five Law & Order series. Unfortunately, the output is partially erased and replaced with blanks.

```{r}
#| results: hide
#| echo: true
lo_titles <- c("Law & Order", "Law & Order: Criminal Intent", "Law & Order: SVU")

law_and_order |>
  filter(title %in% lo_titles) |>
  group_by(title) |>
  summarize(mean_av_rating = mean(av_rating), sd_av_rating = sd(av_rating))
```

```         
# A tibble: 5 × 3
  title                         mean_av_rating sd_av_rating
  <chr>                                  <dbl>        <dbl>
1 Law & Order                            _(1)_        0.106
2 Law & Order: Criminal Intent            8.20        0.129
4 Law & Order: SVU                        8.67        _(2)_
```

Based on the visualizations you've seen of these data so far, which of the following is [**true**]{.underline} about the blanks in the output? **Select all that are true.**

a.  The **mean** of average ratings (Blank 1) of Law & Order seasons is **lower** than the other two means.

b.  The **mean** of average ratings (Blank 1) of Law & Order seasons is **higher** than the other two means.

c.  The **standard deviation** of average ratings of Law & Order: SVU seasons (Blank 2) is **lower** than the other two standard deviations.

d.  The **standard deviation** of average ratings of Law & Order: SVU seasons (Blank 2) is **higher** than the other two standard deviations.

e.  The **standard deviation** of average ratings of Law & Order: SVU seasons (Blank 2) is **between** the other two standard deviations.

```{r}
#| label: blizzard-data-prep
blizzard_salary <- blizzard_salary |>
  mutate(
    annual_salary = case_when(
      salary_type == "week" ~ current_salary * 52,
      salary_type == "hour" ~ current_salary * 40 * 52,
      TRUE ~ current_salary
    ),
    performance_rating = if_else(
      performance_rating == "Developing",
      "Poor",
      performance_rating
    )
  ) |>
  filter(salary_type != "week") |>
  mutate(
    salary_type = if_else(salary_type == "hour", "Hourly", "Salaried")
  ) |>
  filter(!is.na(annual_salary)) |>
  select(percent_incr, salary_type, annual_salary, performance_rating)
```

## Blizzard salaries

In 2020, employees of Blizzard Entertainment circulated a spreadsheet to anonymously share salaries and recent pay increases amidst rising tension in the video game industry over wage disparities and executive compensation.
(Source: [Blizzard Workers Share Salaries in Revolt Over Pay](https://www.bloomberg.com/news/articles/2020-08-03/blizzard-workers-share-salaries-in-revolt-over-wage-disparities))

The name of the data frame used for this analysis is `blizzard_salary` and the variables are:

-   `percent_incr`: Raise given in July 2020, as percent increase with values ranging from 1 (1% increase) to 21.5 (21.5% increase)

-   `salary_type`: Type of salary, with levels `Hourly` and `Salaried`

-   `annual_salary`: Annual salary, in USD, with values ranging from \$50,939 to \$216,856.

-   `performance_rating`: Most recent review performance rating, with levels `Poor`, `Successful`, `High`, and `Top`.
    The `Poor` level is the lowest rating and the `Top` level is the highest rating.

The first ten rows of `blizzard_salary` are shown below:

```{r}
blizzard_salary |>
  select(percent_incr, salary_type, annual_salary, performance_rating)
```

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

@fig-blizzard-hist-1 and @fig-blizzard-hist-2 show the distributions of annual salaries of hourly and salaried workers.
The two figures show the same data, with the facets organized across rows and across columns.
Which of the two figures is better for comparing the median annual salaries of hourly and salaried workers.
Explain your reasoning.

::: {#fig-blizzard-hist}
```{r}
#| label: fig-blizzard-hist-1
#| fig-cap: Option 1
#| fig-width: 8
#| fig-asp: 0.5
ggplot(blizzard_salary, aes(x = annual_salary)) +
  geom_histogram(binwidth = 10000, color = "white") +
  facet_wrap(~salary_type, ncol = 1) +
  scale_x_continuous(labels = label_dollar(scale = 1/1000, suffix = "K"), breaks = seq(0, 250000, 25000)) +
  labs(
    x = "Annual salary (in USD)",
    y = "Count"
  )
```

```{r}
#| label: fig-blizzard-hist-2
#| fig-cap: Option 2
#| fig-width: 8
#| fig-asp: 0.3

ggplot(blizzard_salary, aes(x = annual_salary)) +
  geom_histogram(binwidth = 10000, color = "white") +
  facet_wrap(~salary_type, nrow = 1) +
  scale_x_continuous(labels = label_dollar(scale = 1/1000, suffix = "K"), breaks = seq(0, 250000, 25000)) +
  labs(
    x = "Annual salary (in USD)",
    y = "Count"
  )
```

Distribution of annual salaries of Blizzard employees
:::

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Suppose your teammate wrote the following code as part of their analysis of the data.

```{r}
#| label: blizzard-summarize
#| results: false
#| echo: true
blizzard_salary |>
  group_by(salary_type) |>
  summarize(
    mean_annual_salary = mean(annual_salary),
    median_annual_salary = median(annual_salary)
  )
```

They then printed out the results shown below.
Unfortunately one of the numbers got erased from the printout.
It's indicated with `_____` below.

```         
# A tibble: 2 × 3
  salary_type mean_annual_salary median_annual_salary
  <chr>                    <dbl>                <dbl>
1 Hourly                  63003.               54246.
2 Salaried                90183.               _____
```

Which of the following is the best estimate for that erased value?

a.  30,000

b.  50,000

c.  80,000

d.  100,000

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Which distribution of annual salaries has a higher standard deviation?

a.  Hourly workers

b.  Salaried workers

c.  Roughly the same

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Which of the following alternate plots would also be useful for visualizing the distributions of annual salaries of hourly and salaried workers?
Choose all that apply.

a\.
Box plots

b\.
Density plots

c\.
Pie charts

d\.
Waffle charts

e\.
Histograms

f\.
Scatterplots

{{< pagebreak >}}

```{r}
question0 <- q_no + 1
q_no <- q_no + 3
```

### Questions `{r} question0` - `{r} q_no`

Another teammate creates the following two plots.

```{r}
#| label: fig-salary-performance-rating
#| layout-ncol: 2
#| fig-cap: Distribution of salary type by performance rating
#| fig-width: 5
#| fig-asp: 0.5
#| fig-subcap: 
#|   - Option 1
#|   - Option 2

blizzard_salary |>
  drop_na(salary_type, performance_rating) |>
  ggplot(aes(x = salary_type, fill = fct_rev(performance_rating))) +
  geom_bar() +
  scale_fill_viridis_d() +
  labs(
    x = "Salary type",
    y = "Count",
    fill = "Performance\nrating"
  )

blizzard_salary |>
  drop_na(salary_type, performance_rating) |>
  ggplot(aes(x = salary_type, fill = fct_rev(performance_rating))) +
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(
    x = "Salary type",
    y = "Proportion",
    fill = "Performance\nrating"
  )
```

**Question `{r} question0`:** Your teammate asks you for help deciding which one to use in the final report for visualizing the relationship between performance rating and salary type.
In 1-3 sentences, can you help them make a decision, justify your choice, and write the narrative that should go with the plot?

::: {.content-visible when-format="pdf"}
$\vspace{2cm}$
:::

**Question `{r} question0 + 1`:** A friend with a keen eye points out that the number of observations in @fig-salary-performance-rating-1 seems lower than the total number of observations in `blizzard_salary`.
What might be going on here?
Explain your reasoning.

::: {.content-visible when-format="pdf"}
$\vspace{2cm}$

{{< pagebreak >}}
:::

**Question `{r} question0 + 2`:** Below are the proportions of performance ratings for hourly and salaried workers.
Place these values in the corresponding segments in @fig-salary-performance-rating-2.

```{r}
#| label: salary-type-perfecormance-table

blizzard_salary |>
  drop_na(salary_type, performance_rating) |>
  count(salary_type, performance_rating) |>
  pivot_wider(names_from = salary_type, values_from = n, values_fill = 0) |>
  mutate(across(-performance_rating, ~ .x / sum(.x)))
```

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

You're reviewing another team's work and they made the following visualization:

```{r}
#| warning: false
#| fig-width: 10
#| fig-asp: 0.5

ggplot(blizzard_salary, aes(x = annual_salary, percent_incr)) +
  geom_point() +
  facet_wrap(~performance_rating)
```

And they wrote the following interpretation for the relationship between annual salary and percent increase for Top performers:

> The relationship is positive, having a higher salary results in a higher percent increase.
> There is one clear outlier.

Which of the following are the most accurate and helpful peer review notes for this interpretation? Choose all that apply.

a.  The interpretation is complete and perfect, no changes needed!
a.  The interpretation doesn't mention the direction of the relationship.
a.  The interpretation doesn't mention the form of the relationship, which is linear.
a.  The interpretation doesn't mention the strength of the relationship, which is somewhat strong.
a.  There isn't a clear outlier in the plot.
    If any points stand out as potential outliers, more guidance should be given to the reader to identify them (e.g., salary and/or percent increase amount).
a.  The interpretation is causal – we don't know if the cause of the high percent increase is higher annual salary based on observational data.
    The causal direction might be the other way around, or there may be other factors contributing to the apparent relationship.


```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

You fit a model for predicting raises (`percent_incr`) from salaries (`annual_salary`). We'll call this model `raise_1_fit`. A tidy output of the model is shown below.

```{r}
#| label: raise-salary-fit
raise_1_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary, data = blizzard_salary)

tidy(raise_1_fit)
```

Which of the following is the best interpretation of the slope coefficient?

a.  For every additional \$1,000 of annual salary, the model predicts the raise to be higher, on average, by 1.55%.
b.  For every additional \$1,000 of annual salary, the raise goes up by 0.0155%.
c.  For every additional \$1,000 of annual salary, the model predicts the raise to be higher, on average, by 0.0155%.
d.  For every additional \$1,000 of annual salary, the model predicts the raise to be higher, on average, by 1.87%.

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

You then fit a model for predicting raises (`percent_incr`) from salaries (`annual_salary`) and performance ratings (`performance_rating`). We'll call this model `raise_2_fit`. Which of the following is definitely true based on the information you have so far?

a.  Intercept of `raise_2_fit` is higher than intercept of `raise_1_fit`.
b.  Slope of `raise_2_fit` is higher than RMSE of `raise_1_fit`.
c.  Adjusted $R^2$ of `raise_2_fit` is higher than adjusted $R^2$ of `raise_1_fit`.
d.  $R^2$ of `raise_2_fit` is higher $R^2$ of `raise_1_fit`.


```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

The tidy model output for the `raise_2_fit` model you fit is shown below.

```{r}
#| label: raise-salary-rating-fit
raise_2_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary + performance_rating, data = blizzard_salary)

tidy(raise_2_fit)
```

When your teammate sees this model output, they remark 

> "The coefficient for `performance_ratingSuccessful` is negative. That's weird. I guess it means that people who get successful performance ratings get lower raises." 

How would you respond to your teammate?

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Ultimately, your teammate decides they don't like the negative slope coefficients in the model output you created (not that there's anything wrong with negative slope coefficients!), does something else, and comes up with the following model output.

```{r}
blizzard_salary <- blizzard_salary |>
  mutate(
    performance_rating = fct_relevel(
      performance_rating,
      "Poor",
      "Successful",
      "High",
      "Top"
    )
  )

raise_2_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary + performance_rating, data = blizzard_salary)

tidy(raise_2_fit)
```

Unfortunately they didn't write their code in a Quarto document, instead just wrote some code in the Console and then lost track of their work. They remember using the `fct_relevel()` function and doing something like the following:

```{r}
#| eval: false
#| echo: true
blizzard_salary <- blizzard_salary |>
  mutate(performance_rating = fct_relevel(performance_rating, ___))
```

What should they put in the blanks to get the same model output as above?

a.  "Poor", "Successful", "High", "Top"
b.  "Successful", "High", "Top"
c.  "Top", "High", "Successful", "Poor"
d.  Poor, Successful, High, Top

```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Suppose we fit a model to predict `percent_incr` from `annual_salary` and `salary_type`. A tidy output of the model is shown below.

```{r}
#| label: raise-salary-type-fit
raise_3_fit <- linear_reg() |>
  fit(percent_incr ~ annual_salary + salary_type, data = blizzard_salary)

tidy(raise_3_fit)
```

Which of the following visualizations represent this model? Explain your reasoning.

```{r}
#| label: fig-raise-salary-type
#| warning: false
#| layout-ncol: 2
#| fig-cap: |
#|   Visualizations of the relationship between percent increase, annual
#|   salary, and salary type
#| fig-subcap:
#|   - Option 1
#|   - Option 2
#|   - Option 3
#|   - Option 4
ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(aes(shape = salary_type), alpha = 0.5, size = 2) +
  geom_smooth(
    aes(linetype = salary_type),
    method = "lm",
    se = FALSE,
    fullrange = TRUE,
    linewidth = 1.5
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase",
    color = "Salary type",
    linetype = "Salary type",
    shape = "Salary type"
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  theme(legend.position = "top")

ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(aes(shape = salary_type), alpha = 0.5, size = 2) +
  geom_smooth(
    aes(linetype = salary_type),
    se = FALSE,
    fullrange = TRUE,
    linewidth = 1.5
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase"
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  theme(legend.position = "top")

ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(
    aes(shape = salary_type),
    alpha = 0.5,
    size = 2,
    show.legend = FALSE
  ) +
  geom_abline(
    intercept = 1.24,
    slope = 0.0000137,
    color = "#E87d72",
    linewidth = 1.5,
    linetype = "solid"
  ) +
  geom_abline(
    intercept = 1.24 + 0.913,
    slope = 0.0000137,
    color = "#56bcc2",
    linewidth = 1.5,
    linetype = "dashed"
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase",
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1))

ggplot(
  blizzard_salary,
  aes(x = annual_salary, y = percent_incr, color = salary_type)
) +
  geom_point(alpha = 0.5, size = 2, show.legend = FALSE) +
  geom_abline(
    intercept = 1.24,
    slope = 0.0000137,
    color = "#56bcc2",
    linewidth = 1.5,
    linetype = "dashed"
  ) +
  geom_abline(
    intercept = 1.24 + 0.913,
    slope = 0.0000137,
    color = "#E87d72",
    linewidth = 1.5
  ) +
  labs(
    x = "Annual salary",
    y = "Percent increase",
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent(scale = 1))
```



```{r}
q_no <- q_no + 1
```

### Question `{r} q_no`

Suppose you now fit a model to predict the natural log of percent increase, `log(percent_incr)`, from performance rating. The model is called `raise_4_fit`.

```{r}
raise_4_fit <- linear_reg() |>
  fit(log(percent_incr + 0.0001) ~ performance_rating, data = blizzard_salary)
```

You're provided the following:

```{r}
#| echo: true
tidy(raise_4_fit) |>
  select(term, estimate) |>
  mutate(exp_estimate = exp(estimate))
```

Based on this, which of the following is true?

a\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by 10.25% compared to the employees with Poor performance rating.

b\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by 6.93% compared to the employees with Poor performance rating.

c\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by a factor of 1025 compared to the employees with Poor performance rating.

d\. The model predicts that the percentage increase employees with Successful performance get, on average, is higher by a factor of 6.93 compared to the employees with Poor performance rating.



## Movies

The data for this part comes from the Internet Movie Database (IMDB). Specifically, the data are a random sample of movies released between 1980 and 2020.

```{r}
#| label: load-data
#| message: false
movies <- read_csv("data/movies.csv")
```

The name of the data frame used for this analysis is `movies`, and it contains the variables shown in @tbl-data-dictionary.

\setcounter{table}{0}

| Variable          | Description                                                             |
|----------------------|--------------------------------------------------|
| `name`            | name of the movie                                                       |
| `rating`          | rating of the movie (R, PG, etc.)                                       |
| `genre`           | main genre of the movie.                                                |
| `runtime`         | duration of the movie                                                   |
| `year`            | year of release                                                         |
| `release_date`    | release date (YYYY-MM-DD)                                               |
| `release_country` | release country                                                         |
| `score`           | IMDB user rating                                                        |
| `votes`           | number of user votes                                                    |
| `director`        | the director                                                            |
| `writer`          | writer of the movie                                                     |
| `star`            | main actor/actress                                                      |
| `country`         | country of origin                                                       |
| `budget`          | the budget of a movie (some movies don't have this, so it appears as 0) |
| `gross`           | revenue of the movie                                                    |
| `company`         | the production company                                                  |

: Data dictionary for `movies` {#tbl-data-dictionary tbl-colwidths="\[20,80\]"}

The first thirty rows of the `movies` data frame are shown in @tbl-data, with variable types suppressed (since we'll ask about them later).


```{r}
#| echo: false
movies_to_mark <- c("Blue City", "Rang De Basanti", "Winter Sleep")

movies <- movies |>
  mutate(
    mark = if_else(name %in% movies_to_mark, TRUE, FALSE),
    rating = case_when(
      rating == "TV-PG" ~ "PG",
      rating == "Unrated" ~ "Not Rated",
      is.na(rating) ~ "Not Rated",
      .default = rating
    ),
    rating = fct_relevel(rating, "G", "PG", "PG-13", "R", "NC-17", "Not Rated")
  ) |>
  arrange(desc(mark)) |>
  relocate(
    name,
    score,
    runtime,
    genre,
    rating,
    release_country,
    release_date,
    budget,
    gross,
    votes,
    year,
    director,
    writer,
    star,
    company,
    country
  )
```

```{=tex}
\begin{landscape}
```

::: {#tbl-data}
First 30 rows of `movies`, with variable types suppressed.

```{r}
#| echo: false
options(
  dplyr.print_min = 30,
  pillar.min_chars = 13,
  pillar.width = 80,
  pillar.sigfig = 6
)

format(movies |> select(!c(mark)))[-3L] |>
  str_remove_all(" <.*?>") |>
  cat(sep = "\n")

options(
  dplyr.print_min = 10,
  pillar.min_chars = 8,
  pillar.width = 80,
  pillar.sigfig = 3
)
```

:::



### Score vs. runtime

In this part, we fit a model predicting `score` from `runtime` and name it `score_runtime_fit`.

```{r}
#| echo: true
score_runtime_fit <- linear_reg() |>
  fit(score ~ runtime, data = movies)
```


@fig-score-runtime visualizes the relationship between `score` and `runtime` as well as the linear model for predicting `score` from `runtime`. The first three movies in @tbl-data are labeled in the visualization as well. Answer all questions in this part based on @fig-score-runtime.

```{r}
#| label: fig-score-runtime-prep
#| echo: false
movies <- movies |>
  mutate(
    runtime = parse_number(runtime),
    shape1 = case_when(
      name == movies_to_mark[1] ~ "circle",
      name == movies_to_mark[2] ~ "square",
      name == movies_to_mark[3] ~ "triangle",
      .default = "circle"
    ),
    shape2 = case_when(
      name == movies_to_mark[1] ~ "circle open",
      name == movies_to_mark[2] ~ "square open",
      name == movies_to_mark[3] ~ "triangle open",
      .default = NA
    ),
    color = if_else(mark, "black", "gray50"),
    alpha = if_else(mark, 1, 0.5)
  )
```

```{r}
#| label: fig-score-runtime
#| fig-cap: Scatterplot of `score` vs. `runtime` for `movies`.
#| echo: false
#| message: false
#| fig-asp: 0.5
#| fig-width: 6.5
ggplot(movies, aes(x = runtime, y = score)) +
  geom_point(aes(shape = shape1, color = color, alpha = alpha)) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_shape_identity() +
  scale_alpha_identity() +
  scale_color_identity() +
  geom_point(
    data = movies |> filter(mark),
    aes(shape = shape2),
    size = 3
  ) +
  geom_label(
    data = movies |> filter(mark),
    aes(label = name),
    size = 3,
    nudge_x = 1,
    nudge_y = c(-0.4, -0.4, 0.4)
  ) +
  scale_x_continuous(breaks = seq(80, 240, 20))
```


```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Partial code for producing @fig-score-runtime is given below. Which of the following goes in the blank on Line 2? **Select all that apply.**


```{r}
#| eval: false
#| echo: true
#| code-line-numbers: true
movies |>
  mutate(runtime = ___) |>
  ggplot(aes(x = runtime, y = score)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE)
# additional code for annotating Blue City on the plot
```

a.  `grepl(" mins", runtime)`

b.  `grep(" mins", runtime)`

c.  `str_remove(runtime, " mins")`

d.  `as.numeric(str_remove(runtime, " mins"))`

e.  `na.rm(runtime)`

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Based on this model, order the three labeled movies in @fig-score-runtime in decreasing order of the magnitude (absolute value) of their residuals.

a.  Winter Sleep \> Rang De Basanti \> Blue City

b.  Winter Sleep \> Blue City \> Rang De Basanti

c.  Rang De Basanti \> Winter Sleep \> Blue City

d.  Blue City \> Winter Sleep \> Rang De Basanti

e.  Blue City \> Rang De Basanti \> Winter Sleep



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

```{r}
#| include: false
score_runtime_rsq <- round(glance(score_runtime_fit)$r.squared * 100, 0)
```

The R-squared for the model visualized in @fig-score-runtime is `{r} score_runtime_rsq`%. Which of the following is the [**best**]{.underline} interpretation of this value?

a.  `{r} score_runtime_rsq`% of the variability in movie runtimes is explained by their scores.

b.  `{r} score_runtime_rsq`% of the variability in movie scores is explained by their runtime.

c.  The model accurately predicts scores of `{r} score_runtime_rsq`% of the movies in this sample.

d.  The model accurately predicts scores of `{r} score_runtime_rsq`% of all movies.

e.  The correlation between scores and runtimes of movies is `{r} score_runtime_rsq / 100`.

### Score vs. runtime or year

The visualizations below show the relationship between `score` and `runtime` as well as `score` and `year`, respectively. Additionally, the lines of best fit are overlaid on the visualizations.

```{r}
#| layout-ncol: 2
#| echo: false
#| message: false
ggplot(movies, aes(x = runtime, y = score)) +
  geom_point(alpha = 0.5, color = "gray40") +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 1.2)

ggplot(movies, aes(x = year, y = score)) +
  geom_point(alpha = 0.5, color = "gray40") +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 1.2)
```

The correlation coefficients of these relationships are calculated below, though some of the code and the output are missing. Answer all questions in this part based on the code and output shown below.

```{r}
#| include: false
movies |>
  summarize(
    r_score_runtime = cor(runtime, score),
    r_score_year = cor(year, score)
  )
```

```{r}
#| eval: false
#| echo: true
movies |>
  __blank_1__(
    r_score_runtime = cor(runtime, score),
    r_score_year = cor(year, score)
  )
```

```         
# A tibble: 1 × 2
  r_score_runtime r_score_year
            <dbl>        <dbl>
1           0.434. __blank_2__       
```

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Which of the following goes in `__blank_1__`?

a.  `summarize`

b.  `mutate`

c.  `group_by`

d.  `arrange`

e.  `filter`

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

What can we say about the value that goes in `__blank_2__`?

a.  `NA`

b.  A value between 0 and 0.434.

c.  A value between 0.434 and 1.

d.  A value between 0 and -0.434.

e.  A value between -1 and -0.434.


### Score vs. runtime and rating

In this part, we fit a model predicting `score` from `runtime` and `rating` (categorized as G, PG, PG-13, R, NC-17, and Not Rated), and name it `score_runtime_rating_fit`.

```{r}
score_runtime_rating_fit <- linear_reg() |>
  fit(score ~ runtime + rating, data = movies)
```

The model output for `score_runtime_rating_fit` is shown in @tbl-score-runtime-rating-tidy. Answer all questions in this part based on @tbl-score-runtime-rating-tidy.

```{r}
#| label: tbl-score-runtime-rating-tidy
#| tbl-cap: Regression output for `score_runtime_rating_fit`.
#| echo: false
tidy(score_runtime_rating_fit) |>
  knitr::kable(digits = 3)
```



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Which of the following is [**TRUE**]{.underline} about the intercept of `score_runtime_rating_fit`? **Select all that are true.**

a.  Keeping runtime constant, G-rated movies are predicted to score, on average, 4.525 points.

b.  Keeping runtime constant, movies without a rating are predicted to score, on average, 4.525 points.

c.  Movies without a rating that are 0 minutes in length are predicted to score, on average, 4.525 points.

d.  All else held constant, movies that are 0 minutes in length are predicted to score, on average, 4.525 points.

e.  G-rated movies that are 0 minutes in length are predicted to score, on average, 4.525 points.



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Which of the following is the [**best**]{.underline} interpretation of the slope of `runtime` in `score_runtime_rating_fit`?

a.  All else held constant, as runtime increases by 1 minute, the score of the movie increases by 0.021 points.

b.  For G-rated movies, all else held constant, as runtime increases by 1 minute, we predict the score of the movie to be higher by 0.021 points on average.

c.  All else held constant, for each additional minute of runtime, movie scores will be higher by 0.021 points on average.

d.  G-rated movies that are 0 minutes in length are predicted to score 0.021 points on average.

e.  For each higher level of rating, the movie scores go up by 0.021 points on average.

```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

Fill in the blank:

> R-squared for `score_runtime_rating_fit` (the model predicting `score` from `runtime` and `rating`) \_\_\_\_\_\_\_\_\_ the R-squared the model `score_runtime_fit` (for predicting `score` from `runtime` alone).

a.  is less than

b.  is equal to

c.  is greater than

d.  cannot be compared (based on the information provided) to

e.  is both greater than and less than



```{r}
q_no <- q_no + 1
```

#### Question `{r} q_no`

The model `score_runtime_rating_fit` (the model predicting `score` from `runtime` and `rating`) can be visualized as parallel lines for each level of `rating`. Which of the following is the equation of the line for R-rated movies?

a.  $\widehat{score} = (4.525 - 0.257) + 0.021 \times runtime$

b.  $score = (4.525 - 0.257) + 0.021 \times runtime$

c.  $\widehat{score} = 4.525 + (0.021 - 0.257) \times runtime$

d.  $score = 4.525 + (0.021 - 0.257) \times runtime$

e.  $\widehat{score} = (4.525 + 0.021) - 0.257 \times runtime$



