---
title: "AE 05: Diamonds!"
format: html
---

We want to understand how "The 4 Cs" predict a diamond's price. We have a dataset with 5,000 rows, where each row is a diamond, and along the columns we record each diamond's...

- `price`: in USD;
- `carat`: weight of the diamond;
- `cut`: quality of the cut (Fair, Good, Very Good, Premium, Ideal);
- `color`: diamond colour, from J (worst) to D (best);
- `clarity`: a measurement of how clear the diamond is;
    - I (worst), SI, VS, VVS, IF (best);
- `length_mm`;
- `width_mm`.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
```

## Load and prepare the data

```{r}
#| label: load-data
#| message: false
diamonds <- read_csv("data/diamonds.csv")
```

These data have several *categorical* variables: `cut`, `color`, and `clarity`. By default these variables will have their levels ordered alphabetically:

```{r}
#| label: level-ordering

diamonds <- diamonds |>
  mutate(
    cut = factor(cut),
    clarity = factor(clarity),
    color = factor(color)
  )

levels(diamonds$cut)
levels(diamonds$clarity)
levels(diamonds$color)
```

Because these levels do have a natural ordering corresponding to the increasing level of quality, we should force the computer to respect that:

```{r}
#| label: re-level

diamonds <- diamonds |>
  mutate(
    cut = fct_relevel(cut, "Fair", "Good", "Very Good", "Premium", "Ideal"),
    clarity = fct_relevel(clarity, "I", "SI", "VS", "VVS", "IF")
  )

levels(diamonds$cut)
levels(diamonds$clarity)
```

## Exploratory analysis

Visualize the bivariate relationship between `price` and each of the 4 Cs.

```{r}
#| label: price-carat

ggplot(diamonds, aes(x = carat, y = price)) + 
  geom_point()
```

```{r}
#| label: price-cut

ggplot(diamonds, aes(x = cut, y = price)) + 
  geom_boxplot()
```

```{r}
#| label: price-clarity

ggplot(diamonds, aes(x = clarity, y = price)) + 
  geom_boxplot()
```

```{r}
#| label: price-color

ggplot(diamonds, aes(x = color, y = price)) + 
  geom_boxplot()
```

Put price, carat, and clarity on one plot:

```{r}
#| label: price-carat-clarity

ggplot(diamonds, aes(x = carat, y = price, color = clarity)) + 
  geom_point()
```

## Empty model

Fit the model for `price` that has no predictors:

```{r}
#| label: empty-model

empty_fit <- linear_reg() |>
  fit(price ~ 1, diamonds)

tidy(empty_fit)
```

Interpret the coefficient estimate:

- The average price of a diamond is $3955.28.

## Simple linear regression

Fit the model that predicts `price` from `carat`:

```{r}
#| label: price-carat-model

price_carat_fit <- linear_reg() |>
  fit(price ~ carat, diamonds)

tidy(price_carat_fit)
```

Write the equation for the equation of the fitted model:

$$
\widehat{price}=-2360.58 + 7916.132\times carat.
$$

Interpret the coefficient estimates:

- A 0-carat diamond is predicted to cost $-2360.58 on average. You can't have a 0-carat diamond, and the price won't be negative, so this is doubly meaningless;
- A 1-carat increase predicts an average increase in price of $7916.13.

## Additive model

Fit the additive model that predicts `price` from `carat` and `clarity`:

```{r}
#| label: price-carat-clarity-model-1

price_carat_clarity_fit <- linear_reg() |>
  fit(price ~ carat + clarity, diamonds)

tidy(price_carat_clarity_fit)
```

$$
\widehat{price}=-6989.76 + 8465.86\times carat + 3474.28\times SI + 4570.70\times VS + 5214.40\times VVS + 5515.78\times IF
$$

Interpret some coefficient estimates:

- A 0-carat *I* diamond is predicted to cost $-6989.76 on average;
- Holding clarity constant, a 1-carat increase predicts an average increase in price of $8465.86. That holds for every type of diamond, because the lines are parallel;
- Holding carat constant, an *SI* diamond is expected to cost $3474.28 more than an *I* diamond, on average;
- Holding carat constant, an *VS* diamond is expected to cost $4570.70 more than an *I* diamond, on average;
- Holding carat constant, an *VSS* diamond is expected to cost $5214.40 more than an *I* diamond, on average;
- Holding carat constant, an *IF* diamond is expected to cost $5515.78 more than an *I* diamond, on average.

## Interaction model

Fit the interaction model that predicts `price` from `carat` and `clarity`:


```{r}
#| label: price-carat-clarity-model-2

price_carat_clarity_int_fit <- linear_reg() |>
  fit(price ~ carat * clarity, diamonds)

tidy(price_carat_clarity_int_fit)
```

$$
\begin{aligned}
\widehat{price}
=
-1543.06 + 4148.06\times carat &-1471.07\times SI -1202.31\times VS -1084.20\times VVS -1610.30\times IF
\\
&+ 3782.50\times carat*SI \\
&+ 4748.43\times carat*VS \\
&+ 5855.87\times carat*VVS \\
&+ 7626.83\times carat*IF.
\end{aligned}
$$

Interpret some coefficient estimates:

- A 0-carat *I* diamond is predicted to cost $-1543.06 on average;
- For an *I* diamond, a 1-carat increase predicts an average increase in price of $4148.06;
- A 0-carat *SI* diamond is predicted to cost $1471.07 *less* than a 0-carat *I* diamond, on average;
- A 0-carat *VS* diamond is predicted to cost $1202.31 *less* than a 0-carat *I* diamond, on average;
- A 0-carat *VVS* diamond is predicted to cost $1084.2 *less* than a 0-carat *I* diamond, on average;
- A 0-carat *IF* diamond is predicted to cost $1610.3 *less* than a 0-carat *I* diamond, on average;
- A 1-carat increase predicts an average price change for *SI* diamonds that is $3782.5 *greater* than for *I* diamonds;
- A 1-carat increase predicts an average price change for *VS* diamonds that is $4748.43 *greater* than for *I* diamonds;
- A 1-carat increase predicts an average price change for *VVS* diamonds that is $5855.87 *greater* than for *I* diamonds;
- A 1-carat increase predicts an average price change for *IF* diamonds that is $7626.83 *greater* than for *I* diamonds.

## "Full" model

Fit the additive model that predicts `price` from all other variables in the dataset:


```{r}
#| label: full-model

full_model <- linear_reg() |>
  fit(price ~ ., diamonds)

tidy(full_model)
```

## Selection 

Use adjusted $R^2$ to select the "best" model for `price`.

```{r}
#| label: selection

empty_fit |> glance() |> pull(adj.r.squared)

price_carat_fit |> glance() |> pull(adj.r.squared)

price_carat_clarity_fit |> glance() |> pull(adj.r.squared)

price_carat_clarity_int_fit |> glance() |> pull(adj.r.squared)

full_model |> glance() |> pull(adj.r.squared)
```

The interaction model is "best."


